Modeling potential JCPOA breakout scenarios using Cyclus

### 2018-01-27 paper.tex
small changes and review of entire paper

The Joint Comprehensive Plan of Action (JCPOA)\cite{jcpoa} %cite JCPOA agreement
limits the Iranian nuclear program and, in return, lifts many of the economic
sanctions that had been placed on Iran. Under the JCPOA, Iran is severely
limited in its ability to enrich Uranium. This paper presents a model to probe
the enriching ability of various cascade configurations that are related to
JCPOA agreement limits. This analysis was performed using the Cyclus fuel cycle
simulator\cite{cyclus} %cite Fundamental concepts in the Cyclus nuclear fuel cycle
simulation framework in order to demonstrate that such analysis can be
integrated into a fuel cycle with material flowing among multiple facilities.

The Cyclus fuel cycle simulator is designed to support the dynamic addition
novel facility models into fuel cycles both declared and undeclared facilities.
The CascadeEnrich archetype\cite{mbmore.2018} was developed to model gas centrifuge
enrichment cascades at a higher fidelity than previous enrichment models,
relying on individual centrifuge and cascade design parameters to build an ideal
cascade configuration instead of integral facility parameters. Once a
CascadeEnrich facility is deployed, it may be asked to participate in the fuel
cycle at a variety of flow rates and enrichments over time, constrained by the
initial design of the cascade.

@@

The cascade is built to be ideal, defined by $\alpha =\beta = const$ for all
stage of the cascade, where $\alpha$ and $\beta$ respectively represent the feed
to product and the feed to tail enrichment factors.  $\alpha$ can be expressed
as a function of the feed rate $F$, the separative performance $\delta
U(\theta)$ and $\theta$, and $\beta$ as a function of $alpha$ and $\theta$:

@@

From equation \eqref{eq:alpha} and \eqref{eq:beta} it is possible to determine
the cut, or the ratio of product flow to feed flow.
%(all other centrifuge parameters been fixed) allowing the have $\alpha=\beta$
required to build an ideal cascade.

Since $\alpha_{i}$ and $\beta_{i}$ remain constant, only the value of the cut,
$\theta_{i}$, changes in each stage of a cascade.

@@

This algorithm assumes that the corresponding separative power $\delta U$ (not
re-computed) can be achieved with the chosen centrifuge design. Once
$\theta_{i}$ determined, it is possible to compute the product and the tail
assay.

The design of the cascade is performed through 2 steps. First one determines the
configuration and number of stages, adding stages following the previously
described procedure, until $N'_{last} > N'_{requested}$ and $N''_{last} <
N''_{requested}$. (Note that $N'_{last}$ corresponds to the product assay of the
last enrichment stage, and $N''_{last}$ to the tail assay of the last stripping
stage). This allows to determine the number of enriching and stripping stages as
well as their enrichment properties ($N_{i}$, $N'_{i}$,
$N''_{i}$,$\theta_{i}$i).
% might be more useful to the reader to say this in words instead of variables,
% i.e. adding stages until the product assay of the final stage is greater than
% or equal to than the desired assay, and the tails assay is similarly less than
% or equal to the desired tails assay

The second step determines how to populate the cascade with the user-defined
maximum number of centrifuges. One first solves the linear flow equation to
determine the theoretical flow in the cascade.
%% 
%% \begin{strip}
%% \begin{equation}
%% \setcounter{MaxMatrixCols}{20}
%% \begin{bmatrix}
%% -1         & 1-\theta_{S+1} & 0              & 0 & ... & 0 & 0            & 0   & 0              & 0 & ... & 0            & 0            & 0 \\
%% \theta_{S} & -1             & 1-\theta_{S+2} & 0 & ... & 0 & 0            & 0   & 0              & 0 & ... & 0            & 0            & 0 \\
%%            &                &                &   &     &   &              & ... &                &   &     &              &              &   \\
%% 0          & 0              & 0              & 0 & ... & 0 & \theta_{-1}  & -1  & 1 - \theta_{1} & 0 & ... & 0            & 0            & 0 \\
%%            &                &                &   &     &   &              & ... &                &   &     &              &              &   \\
%% 0          & 0              & 0              & 0 & ... & 0 & 0            & 0   & 0              & 0 & ... & \theta_{E-2} & -1           & 1-\theta_{E} \\
%% 0          & 0              & 0              & 0 & ... & 0 & 0            & 0   & 0              & 0 & ... & 0            & \theta_{E-1} & -1
%% \end{bmatrix}
%% \times
%% \begin{bmatrix}
%%      F_{S}   \\
%%      F_{S+1} \\
%%      \cdots  \\
%%      F_{0}   \\
%%     \cdots   \\
%%     F_{E-1}  \\
%%     F_{E}
%% \end{bmatrix}
%% =
%% \begin{bmatrix}
%%      0   \\
%%      0 \\
%%      \cdots  \\
%%      F   \\
%%     \cdots   \\
%%     0  \\
%%     0
%% \end{bmatrix}
%% %\caption{caption needed!}
%% \label{eq:flow}
%% \end{equation}
%% \end{strip}
Once the relative flow of each stage has been determined, the
cascade can be populated with actual machines up the stages
until either the maximum number available of machines or the maximum feed
flow is reached.

@@

The tuning method outlined here does not re-optimize $\theta_i$ based on the
true flow enrichment. As $\delta U$ and $\alpha$ do not depend on the stage feed
assay $N'$ and the feed to product ratio $\alpha$ do not change from stage to
stage. According to equation \eqref{eq:beta}, when $\alpha$ and $\theta$ are
fixed, if the feed assay, $N$ changes, then $\beta$ will change accordingly.
This brakes the ideal status of the cascade, i.e.  $N_{i} \neq N'_{i-1} \neq
N''_{i+1}$.

In order to compute the proper product and tails assay at each stages\, the
tails and the product from respectively the next and the previous stage must be
blended in order to determine the correct stage feed assay. As this is a obvious
cycling problem, it has been chosen to solve it iteratively: all feed assay are
iteratively updated, blending the proper product and tail, then using the
updated feed assay, the new product and tail assay are recomputed. This process
is repeated until the change in assays is smaller than the set precision (1e-8
by default).

@@

As explained previously, the goal of this work is to assess the rate of the
highly-enriched uranium (HEU) production using enrichment cascades designed
within the JCPOA agreement constraints. For this work, two different cases will
be investigated and compared to the reference case.

All cases will be considered with and without tail recycling. Figure
\ref{fig:flow} shows the different flows between the different facilities in the
fleet, in a configuration with two different levels of cascade without tail
recycling. When considering the tail recycling, all the flow going from a
cascade to the waste, are going from the cascade $i$ to the cascade $i-1$.  The
Storage are not required in Cyclus they have been added to ensure the proper
material flow between the facilities (if a receiver facility input buffer is
full then the sender facility will stop sending material to it, and will stop
working when its own output buffer will be filled).

@@

The reference case corresponds to the most ideal cascade design--utilizing all
of the 5060 centrifuges available within JCPOA to produce an enriched product at
$>90\%$ of U-235, stripping the tail product up to 0.28 $\%$ from a feed of
natural uranium.  The design cascade in this case, includes 4 stripping stages,
and 38 enriching stages. The Feed/Product/Tail (F/P/T) assays are respectively
0.0071/0.903/0.0028.

@@

The case 1 is designed to configure 30 cascades in the scenario (the maximum
allowable under the JCPOA agreement), all designed the same way. The basic
properties are: less than 169 centrifuges per cascade, F/P/T assays of
0.0071/0.035/0.003. The corresponding design includes 167 centrifuges, with
respective F/P/T assays of 0.0071/0.0412842/0.00290548.

@@

Case 2 also limited to 30 cascades in the scenario (the maximum allowable under
the JCPOA agreement), but this time each level of cascades are designed to take
feed from the previous level's product. Each cascade has 4 stripping stages and
10 enriching stages.

\begin{table}[htb]
\centering
\begin{tabular}{cllll}
\toprule

@@

\bottomrule
\end{tabular}
  \caption{Summary of cascade properties for each level}
  \label{tab:cascadelvl}
\end{table}

@@

A last round of optimization if then ran in order to limit the number of cascade
level to the amount of level required produce HEU, and then to populate the
different level with a total of 30 cascades in order to maximize the HEU
production rate.

@@

As illustrated in Table \ref{tab:cascade_theta}, the effective cut
of each level increases up to $0.54$ for the fifth stage for case 2. The high
effective cut value increase the product production rate but decrease the
enrichment ratio.
\begin{table}[htb]
\centering
\begin{tabular}{cll}
\toprule

@@

\bottomrule
\end{tabular}
  \caption{Cascade effective $\theta$ by the level and case.}
  \label{tab:cascade_theta}
\end{table}

\subsubsection{tails recycling}
When recycling the tails from the different cascade levels, both cases behave
the same way. As the tails have an higher enrichment than the product then are
blended to, recycling the tail has the effect to bump up the enrichment of the
product at each level (see Figure \ref{fig:assay_c1_r} and \ref{fig:assay_c2_r}).

@@

Note that for the case 2, where the fourth stage could appear as useless only the
blending of its tail allows the product of the third one to pass the $90\%$ threshold.

@@

\subsection{HEU flow rate}
As shown Figure \ref{fig:heu_prod}, if using all the centrifuges available to
build a big cascade dedicated to HEU production, only 8 months are required to
produce 1 Significant Quantity (SQ) of HEU. Also it take significantly longer to
build 1 SQ of HEU, without recycling the tails respectively 13 years and 3.5 year
for the case 1 and 2 versus about 2 years with tail recycling.

@@

It is also interesting to note that despite requiring more level of cascade to
reach an enrichment of $90\%$ the second case with tail recycling produces HEU
at a equivalent rate than the case 1. The production rate seems to be faster but
require more time to reach the maximum production rate (the material have to
flow through more levels to reach the final level).

>>>>

The Joint Comprehensive Plan of Action (JCPOA)\cite{jcpoa} %cite JCPOA agreement
limits the Iranian nuclear program and, in return, lifts many of the economic
sanctions that had been placed on Iran. Under the JCPOA, Iran is severely
limited in its ability to enrich uranium. This paper presents a model to probe
the enriching ability of various cascade configurations that are related to
JCPOA agreement limits. The analysis was performed using the Cyclus fuel cycle
\cite{cyclus} %cite Fundamental concepts in the Cyclus nuclear fuel cycle
simulation framework in order to demonstrate that such analysis can be
integrated into a fuel cycle with material flowing among multiple facilities.

The Cyclus fuel cycle simulator is designed to support the dynamic addition of
novel facility models into fuel cycles with both declared and undeclared facilities.
The CascadeEnrich archetype\cite{mbmore.2018} was developed to model gas centrifuge
enrichment cascades at a higher fidelity than previous enrichment models,
relying on individual centrifuge and cascade design parameters to build an ideal
cascade configuration instead of integral facility parameters. Once a
CascadeEnrich facility is deployed, it may be asked to participate in the fuel
cycle at a variety of flow rates and enrichments over time, constrained by the
initial design of the cascade.

@@

The cascade is built to be ideal, defined by $\alpha =\beta = const$ for all
stage of the cascade, where $\alpha$ and $\beta$ respectively represent the feed
to product and the feed to tail enrichment factors.  $\alpha$ can be expressed
as a function of the feed rate $F$, the separative performance $\delta
U(\theta)$, the cut $\theta$, and $\beta$ as a function of $alpha$ and $\theta$:

@@

From equation \eqref{eq:alpha} and \eqref{eq:beta} it is possible to determine
the cut, or the ratio of product flow to feed flow
%(all other centrifuge parameters been fixed) allowing the have $\alpha=\beta$
required to build an ideal cascade.

Since $\alpha_{i}$ and $\beta_{i}$ remain constant, only the value of the cut,
$\theta_{i}$, changes in each stage $i$ of a cascade.

@@

This algorithm assumes that the corresponding separative power $\delta U$ (not
re-computed) can be achieved with the chosen centrifuge design. Once
$\theta_{i}$ is determined, it is possible to compute the product and the tail
assay.

The design of the cascade is performed through 2 steps. First one determines the
configuration and number of stages, adding stages until the product assay of the final stage is greater than or equal to than the desired assay, and the tails assay is similarly less than or equal to the desired tails assay. This determines the number of enriching and stripping stages as
well as their enrichment properties ($N_{i}$, $N'_{i}$,
$N''_{i}$,$\theta_{i}$i).

The second step determines how to populate the cascade with the user-defined
maximum number of centrifuges. One then solves the linear flow equation to
determine the theoretical flow in the cascade.
%%
%% \begin{strip}
%% \begin{equation}
%% \setcounter{MaxMatrixCols}{20}
%% \begin{bmatrix}
%% -1         & 1-\theta_{S+1} & 0              & 0 & ... & 0 & 0            & 0   & 0              & 0 & ... & 0            & 0            & 0 \\
%% \theta_{S} & -1             & 1-\theta_{S+2} & 0 & ... & 0 & 0            & 0   & 0              & 0 & ... & 0            & 0            & 0 \\
%%            &                &                &   &     &   &              & ... &                &   &     &              &              &   \\
%% 0          & 0              & 0              & 0 & ... & 0 & \theta_{-1}  & -1  & 1 - \theta_{1} & 0 & ... & 0            & 0            & 0 \\
%%            &                &                &   &     &   &              & ... &                &   &     &              &              &   \\
%% 0          & 0              & 0              & 0 & ... & 0 & 0            & 0   & 0              & 0 & ... & \theta_{E-2} & -1           & 1-\theta_{E} \\
%% 0          & 0              & 0              & 0 & ... & 0 & 0            & 0   & 0              & 0 & ... & 0            & \theta_{E-1} & -1
%% \end{bmatrix}
%% \times
%% \begin{bmatrix}
%%      F_{S}   \\
%%      F_{S+1} \\
%%      \cdots  \\
%%      F_{0}   \\
%%     \cdots   \\
%%     F_{E-1}  \\
%%     F_{E}
%% \end{bmatrix}
%% =
%% \begin{bmatrix}
%%      0   \\
%%      0 \\
%%      \cdots  \\
%%      F   \\
%%     \cdots   \\
%%     0  \\
%%     0
%% \end{bmatrix}
%% %\caption{caption needed!}
%% \label{eq:flow}
%% \end{equation}
%% \end{strip}
Once the relative flow of each stage has been determined, the
cascade can be populated with actual machines up the stages
until either the maximum number available of machines or the maximum feed
flow is reached.

@@

The tuning method outlined here does not re-optimize $\theta_i$ based on the
true flow enrichment. As $\delta U$ and $\alpha$ do not depend on the stage feed
assay ($N'$), they do not change from stage to
stage. According to equation \eqref{eq:beta}, when $\alpha$ and $\theta$ are
fixed, whenthe feed assay ($N$) changes, then $\beta$ will change accordingly.
This breaks the ideal status of the cascade, i.e.  $N_{i} \neq N'_{i-1} \neq
N''_{i+1}$.

In order to compute the proper product and tails assay at each stage, the
tails and the product from respectively the next and the previous stage must be
blended in order to determine the correct stage feed assay. As this is a obvious
cycling problem, an iterative solution has been chosen: all feed assays are
iteratively updated, blending the proper product and tails, then using the
updated feed assay, the new product and tail assays are recomputed. This process
is repeated until the change in assays is smaller than the set precision (1e-8
by default).

@@

As explained previously, the goal of this work is to assess the rate of
highly-enriched uranium (HEU) production using enrichment cascades designed
near, but not within, the JCPOA agreement constraints. For this work, two different cases will
be investigated and compared to a reference case that is based on the most efficient use of the 5060 available centrifuges.

All cases will be considered with and without tail recycling. Figure
\ref{fig:flow} shows the material flows between the different facilities in the
fleet, in a configuration with two "levels" of cascades without tail
recycling. When considering the tail recycling, all of the flow going from a
cascade to the waste, are going from the cascade $i$ to the cascade $i-1$.  The
Storage facilities are not required in Cyclus, but they have been added to ensure the proper
material flow between the facilities.

% you don't need this unless you want to explain how a buffer works (if a receiver facility input buffer is full then the sender facility will stop sending material to it, and will stop working when its own output buffer will be filled).

@@

The reference case corresponds to the most efficient cascade design--utilizing all
of the 5060 centrifuges available within JCPOA in a single large cascade to produce an enriched product at
$>90\%$ of U-235, stripping the tail product up to 0.28 $\%$ from a feed of
natural uranium.  The design cascade in this case, includes 4 stripping stages,
and 38 enriching stages. The Feed/Product/Tail (F/P/T) assays are respectively
0.0071/0.903/0.0028.

@@

The case 1 is designed to configure identical 30 cascades in the scenario (the maximum
allowable under the JCPOA agreement). The basic
properties are: less than 169 centrifuges per cascade, F/P/T assays of
0.0071/0.035/0.003. The corresponding cascade design includes 167 centrifuges, with
respective F/P/T assays of 0.0071/0.0412842/0.00290548.

@@

Case 2 is also limited to 30 cascades in the scenario (the maximum allowable under
the JCPOA agreement), but this time each level of cascades are optimized for the feed they recieve from the previous level's product. Each cascade has 4 stripping stages and
10 enriching stages.

\begin{table}[htb]
\centering
  \caption{Summary of cascade properties for each level}
\begin{tabular}{cllll}
\toprule

@@

\bottomrule
\end{tabular}

  \label{tab:cascadelvl}
\end{table}

@@

One last round of optimization limits the number of cascade
levels to the number of level required produce HEU, and then to populate each level with a total of 30 cascades in order to maximize the HEU production rate.

@@

As illustrated in Table \ref{tab:cascade_theta}, the effective cut
of each level increases up to $0.54$ for the fifth stage for case 2. The high
effective cut value increase the product production rate but decrease the
enrichment ratio.
\begin{table}[htb]
\centering
\caption{Cascade effective $\theta$ by the level and case.}
\begin{tabular}{cll}
\toprule

@@

\bottomrule
\end{tabular}

  \label{tab:cascade_theta}
\end{table}

\subsubsection{Recycling tails}
When recycling the tails from the each cascade level, both cases behave
the same way. As the tails have an higher enrichment than the product they are
blended to, recycling the tails has bumps up the enrichment of the
product at each level (see Figure \ref{fig:assay_c1_r} and \ref{fig:assay_c2_r}).

@@

Note that for case 2, a fourth level is needed while blending the tails for the product of the third level to pass the $90\%$ threshold.

@@

\subsection{HEU flow rate}
As shown Figure \ref{fig:heu_prod}, if all of the available centrifuges are dedicated to HEU production in one massive cascade, only 8 months are required to
produce 1 Significant Quantity (SQ) of HEU. It also takes significantly longer to
stockpile 1 SQ of HEU without recycling the tails, respectively 13 years and 3.5 years
for the case 1 and 2 versus about 2 years with tail recycling.

@@

Despite requiring more levels of cascades to
reach an enrichment of $90\%$, the second case with tail recycling produces HEU
at a equivalent rate to case 1. The production rate is faster, but
require more times to reach the maximum production rate. This is because the material has to
flow through more levels to reach its final level.
